### 2008-10-06 trunk/Papers/rmrs-08/rmrs.tex
Monday update

We define the syntax of \rmrs\ in the style of {\sc clls}
\cite{egg:etal:2001}.  We assume an infinite set of {\em node
  variables} $\nvar=\{X,Y,X_1,\ldots\}$, which we will use as labels,
anchors, and holes; the distinction between these uses will come from
their position in the formula.  Furthermore, we assume an infinite set
of \emph{base variables} $\ovar$, consisting of individual variables
$\{x,x_1,y,\ldots\}$ and event variables $\{e_1,\ldots\}$.  Finally,
we assume a vocabulary of \emph{predicate symbols}
$\pred=\{P,Q,P_1,\ldots\}$, each of which has an \emph{arity} $\ar(P)
\geq 0$.

@@

A node variable $X$ is called a \emph{label} iff $\varphi$ contains an
atom of the form 
$\ep{X}{Y}{P}$ or $Y \qeq X$; $X$ is called an
 a \emph{anchor} in the \rmrs\
$\varphi$ iff $\varphi$ contains an at atom of the form
$\ep{Y}{X}{P}$ or $\ARG_S(X,i)$; and $X$ is
called a \emph{hole} iff 
$\varphi$ contains an atom of the form $\ARG_S(Y,X)$ or $X \qeq Y$.

@@

\rmrs\ as presented here combines similarities to earlier
presentations of \rmrs\ (e.g., \cite{copestake:2003}) and to dominance
constraints \cite{egg:etal:2001}.  We use $\ARG_{\{i\}}$ (with a singleton
set $S$) instead of $\ARG_i$, and $\ARG_\N$ instead of $\ARG_n$.
Furthermore, the elementary predication $\ep{l}{a}{P(v)}$ is a
notational variant of $\{\ep{l}{a}{P}, \ARG_{\{0\}}(l,v)\}$.  In other
words, the syntax defined here differs in some details, but is more
general. The primary differences to dominance constraints are the
distinction of labels and anchors that will separate the labelling and
the parent-child relations, and the use of $\qeq$ instead of
dominance.  Redefining \rmrs\ in the style of dominance constraints
will simplify the formal work below.

@@

\begin{definition}\label{defn:dominance}
$u \qeqdom v$ iff $v = u \cdot 2^k$ for some $k \geq 0$ and all
  symbols $\tau(u), \tau(u \cdot 2), \ldots, \tau(u \cdot 2^{k-1})$
  are quantifiers.  (This is the ``qeq'' relation familiar from the
  \mrs\ and \rmrs\ literature.)

@@

\begin{definition}\label{defn:satisfaction}
Satisfaction of atoms is defined as follows:
$$
\begin{array}{l}
  \tau, \alpha, g, \sigma \models  \ep{X}{Y}{P}
\mbox{ iff}\\
\qquad \tau(\alpha(Y)) \in \sigma(P) \mbox{ and } \alpha(X) \wedgedom
  \alpha(Y) \\
  \tau, \alpha, g, \sigma \models \ARG_S(X,a)
\mbox{ iff exists  $i \in S$ s.t.}\\
\qquad  \alpha(X) \cdot i \in D(\tau) \mbox{ and }
  \tau(\alpha(X) \cdot 
  i) = g(a) \\
  \tau, \alpha, g, \sigma \models \ARG_S(X,Y)
\mbox{ iff  exists $i \in S$ s.t.}\\
\qquad \alpha(X) \cdot i \in D(\tau), \alpha(X) \cdot
  i = \alpha(Y) \mbox{ and }\\
\qquad
s(\alpha(X) \cdot i) = h \\
  \tau, \alpha, g, \sigma \models X \qeq Y
\mbox{ iff } \alpha(X) \qeqdom \alpha(Y) \\
  \tau, \alpha, g, \sigma \models X ={/}\neq Y
\mbox{ iff } \\
\qquad \alpha(X) ={/}\neq \alpha(Y) \\
  \tau, \alpha, g, \sigma \models v_1 ={/}\neq v_2
\mbox{ iff}\\
\qquad g(v_1) ={/}\neq g(v_2) \\
  \tau, \alpha, g, \sigma \models P \spec Q
\mbox{iff $\sigma(P) \subseteq \sigma(Q)$}
\end{array}
$$
A 4-tuple $\tau,\alpha,g,\sigma$ satisfies an \rmrs\ $\varphi$
(written $\tau,\alpha,g,\sigma \models \varphi$) iff it satisfies all
of its elements.
\end{definition}

Notice that one \rmrs\ may be satisfied by multiple trees; we can take
the \rmrs\ to be a partial description of each of these trees.  In
particular, \rmrs s may represent semantic scope ambiguities, missing
information about semantic dependencies, and missing information about
lexical subcategorisation and lexical senses.  For $j=\{1,2\}$,
suppose that $\tau_j,\alpha_j,g_j,\sigma\models \varphi$ and let
$\nvar$ and $\ovar$ be the node and base variables from which
$\varphi$ is built.  Then 
$\varphi$ exhibits a semantic scope ambiguity if there are variables
$Y,Y'\in \nvar$ such that $\alpha_1(Y) \qeqdom \alpha_1(Y')$ and
$\alpha_2(Y')\qeqdom \alpha_2(Y)$.  It exhibits missing information
about semantic dependencies if there are base-language variables $
v,v'\in \ovar$ such that $g_1(v)=g_1(v')$ and $g_2(v)\neq g_2(v')$.
It exhibits missing lexical subcategorisation information if there is
a variable $ Y \in \nvar$ such that $\tau_1(\alpha_1(Y))$ is a
constructor of a different type from $\tau_2(\alpha_2(Y))$  (i.e., the
constructors are of a different arity or and/or their arguments are).
And it exhibits missing lexical sense information
if $\tau_1(\alpha_1(Y))$ and $\tau_2(\alpha_2(Y))$ are different
base-language constructors, but of the same type.

@@

One feature which the above definition of \rmrs\ shares with dominance
constraints is that any satisfiable \rmrs\ has an infinite number of
models which only differ in the areas that the \rmrs\ didn't ``talk
about''.  An example is shown in Fig.~\ref{fig:fat-black-cat}.  This
tree, which represents the meaning of ``every fat black cat chased
some dog'', differs from the model in Fig.~\ref{fig:1} in the subtree
below the restrictor of the universal quantifier.  It is still a model
of (\ref{ex:cat-pos})---all the {\sc ep}s are satisfied.  However, there is
more ``semantic material'' (the ``black'' subformula).  The \rmrs\
didn't specify that this material should be there; but it also didn't
specify that it can't be there, and indeed there is no way in which
the \rmrs\ can enforce that no extra material can be added.

@@

\todo{At this point we should have some examples of solved forms.  The
  running example is a bit lame here because it already is in solved
  form, and therefore only has one solved form.}

>>>>

We define the syntax of \rmrs\ in the style of {\sc clls}
\cite{egg:etal:2001}.  We assume an infinite set of {\em node
  variables} $\nvar=\{X,Y,X_1,\ldots\}$, used as labels,
anchors, and holes; the distinction between these uses will come from
their position in the formula.  Furthermore, we assume an infinite set
of \emph{base variables} $\ovar$, consisting of individual variables
$\{x,x_1,y,\ldots\}$ and event variables $\{e_1,\ldots\}$.  Finally,
we assume a vocabulary of \emph{predicate symbols}
$\pred=\{P,Q,P_1,\ldots\}$.

@@

A node variable $X$ is called a \emph{label} iff $\varphi$ contains an
atom of the form 
$\ep{X}{Y}{P}$ or $Y \qeq X$; $X$ is called an
\emph{anchor} in the \rmrs\
$\varphi$ iff $\varphi$ contains an atom of the form
$\ep{Y}{X}{P}$ or $\ARG_S(X,i)$; and $X$ is
called a \emph{hole} iff 
$\varphi$ contains an atom of the form $\ARG_S(Y,X)$ or $X \qeq Y$.

@@

\rmrs\ as presented here combines similarities to earlier
presentations of \rmrs\ (e.g., \cite{copestake:2003}) and to dominance
constraints \cite{egg:etal:2001}.  We use $\ARG_{\{i\}}$ (with a singleton
set $S$) instead of $\ARG_i$, and $\ARG_\N$ instead of $\ARG_n$.
The elementary predication $\ep{l}{a}{P(v)}$ (see
(\ref{ex:cat-pos})--(\ref{ex:cat-erg})) is a
notational variant of $\{\ep{l}{a}{P}, \ARG_{\{0\}}(a,v)\}$.  In other
words, the syntax defined here differs in some details, but is more
general. The primary differences to dominance constraints are the
distinction of labels and anchors that will separate the labelling and
the parent-child relations, and the use of $\qeq$ instead of
dominance.  Redefining \rmrs\ in the style of dominance constraints
simplifies the formal work below.

@@

\begin{definition}\label{defn:dominance}
$u \qeqdom v$ iff $v = u \cdot 2^k$ for some $k \geq 0$ and all
  symbols $\tau(u), \tau(u \cdot 2), \ldots, \tau(u \cdot 2^{k-1})$
  are quantifiers.  (This is the ``qeq'' relation familiar from
  \newcite{copestake:etal:2005}).

@@

\begin{definition}\label{defn:satisfaction}
Satisfaction of atoms is defined as follows:
$$
\begin{array}{l}
  \tau, \alpha, g, \sigma \models  \ep{X}{Y}{P}
\mbox{ iff}\\
\qquad \tau(\alpha(Y)) \in \sigma(P) \mbox{ and } \alpha(X) \wedgedom
  \alpha(Y) \\
  \tau, \alpha, g, \sigma \models \ARG_S(X,a)
\mbox{ iff exists  $i \in S$ s.t.}\\
\qquad  \alpha(X) \cdot i \in D(\tau) \mbox{ and }
  \tau(\alpha(X) \cdot 
  i) = g(a) \\
  \tau, \alpha, g, \sigma \models \ARG_S(X,Y)
\mbox{ iff  exists $i \in S$ s.t.}\\
\qquad \alpha(X) \cdot i \in D(\tau), \alpha(X) \cdot
  i = \alpha(Y)\\
  \tau, \alpha, g, \sigma \models X \qeq Y
\mbox{ iff } \alpha(X) \qeqdom \alpha(Y) \\
  \tau, \alpha, g, \sigma \models X ={/}\neq Y
\mbox{ iff } \\
\qquad \alpha(X) ={/}\neq \alpha(Y) \\
  \tau, \alpha, g, \sigma \models v_1 ={/}\neq v_2
\mbox{ iff}\\
\qquad g(v_1) ={/}\neq g(v_2) \\
  \tau, \alpha, g, \sigma \models P \spec Q
\mbox{ iff $\sigma(P) \subseteq \sigma(Q)$}
\end{array}
$$
A 4-tuple $\tau,\alpha,g,\sigma$ satisfies an \rmrs\ $\varphi$
(written $\tau,\alpha,g,\sigma \models \varphi$) iff it satisfies all
of its elements.
\end{definition}

Notice that one \rmrs\ may be satisfied by multiple trees; we can take
the \rmrs\ to be a partial description of each of these trees.  In
particular, \rmrs s may represent semantic scope ambiguities, and/or missing
information about semantic dependencies, 
lexical subcategorisation and lexical senses.  For $j=\{1,2\}$,
suppose that $\tau_j,\alpha_j,g_j,\sigma\models \varphi$ and let
$\nvar$ and $\ovar$ be the node and base variables from which
$\varphi$ is built.  Then 
$\varphi$ exhibits a semantic scope ambiguity if there are variables
$Y,Y'\in \nvar$ such that $\alpha_1(Y) \qeqdom \alpha_1(Y')$ and
$\alpha_2(Y')\qeqdom \alpha_2(Y)$.  It exhibits missing information
about semantic dependencies if there are base-language variables
$v,v'\in \ovar$ such that $g_1(v)=g_1(v')$ and $g_2(v)\neq g_2(v')$. 
It exhibits missing lexical subcategorisation information if there is
a $ Y \in \nvar$ such that $\tau_1(\alpha_1(Y))$ is a
constructor of a different type from $\tau_2(\alpha_2(Y))$  (i.e., the
constructors are of a different arity and/or their arguments are).
And it exhibits missing lexical sense information
if $\tau_1(\alpha_1(Y))$ and $\tau_2(\alpha_2(Y))$ are different
base-language constructors, but of the same type.

@@

One feature which the above definition of \rmrs\ shares with dominance
constraints is that any satisfiable \rmrs\ has an infinite number of
models which only differ in the areas that the \rmrs\ didn't ``talk
about''.  An example is shown in Fig.~\ref{fig:fat-black-cat}.  This
tree, which represents the meaning of ``every fat black cat chased
some dog'', is still a model
of (\ref{ex:cat-pos})---all the {\sc ep}s are satisfied.  However, there is
more ``semantic material'' (the ``black'' subformula).  The \rmrs\
didn't specify that this material should be there; but it also didn't
specify that it can't be there, and indeed there is no way in which
the \rmrs\ can enforce that no extra material can be added (e.g.,
consider predicate symbols that are introduced by construction rules
in a deep grammar, which by design are absent from the semantic output
of a {\sc pos} tagger).

@@

\todo{Draw a figure for (\ref{ex:cat-erg}), and explain that the trees
  in Fig 1 are its solved forms.}
